---
title: "Probability exercises"
author: "SMWadgymar"
date: '2022-08-16'
output:
  pdf_document: default
  html_document: default
---

####Exercises

1) Simulating coin flips:  Generate 10 separate random flips with probability 0.7 of producing heads.

```{r}
rbinom(10, 1, .7)
```

Q: What kind of values do you see and what do they represent?

A: 


2) Simulating draws from a binomial distribution:  Generate 100 occurrences of flipping 10 coins, each with 70% probability of producing heads. Produce a table of the probability distribution that you generate (hint: your table should report probabilities and not counts).

```{r}
flips<-rbinom(100, 10, 0.7)

prop.table(table(flips))

#Could also use table(flips)/100 to get the table
```

Q: What kind of values do you produce and what do they represent?  Plot a histogram and describe the shape that your probability distribution takes.  What does the freq=F argument mean?

A: 
```{r}
hist(flips, freq=F)
```



3) Calculate the exact probability that 2 heads will arise from 10 coin flips with a 70% probability of coming up tails.  

```{r}
dbinom(2, 10, 0.3) 
#Note that the word problem lists the probability of getting tails
```

Q: Compare your answer with a simulation of 10,000 trials.  Do the two approaches yield similar results?

A: 
```{r}
mean(rbinom(10000, 10, 0.3)==2)
```



4) Calculate the cumulative probability that at least five coins out of 10 are heads with a 30% probability of coming up heads.  

```{r}
1-pbinom(4, 10, 0.3)
##Note that we have to subtract from 1 to get the right range
```

Q: Compare your answer with a simulation of 10,000 trials.  Do the two approaches yield similar results?

A: 
```{r}
mean(rbinom(10000, 10, 0.3)>= 5)
```



5) Repeat the simulation you ran in exercise (3) where you estimated the exact probability that two heads will occur with 10, 100, 1,000, 10,000, and 100,000 trials.  Produce a plot depicting the number of trails on the x-axis and the associated probabilities you calculated on the y-axis.  

Hint: The c() function lets you combine values into a vector or list.  You will need to make a vector of probabilities for the y axis using and a vector of your sample sizes for the x axis using c(#, #, #, #, #).

```{r}
r10<-mean(rbinom(10, 10, 0.3)==2)
r100<-mean(rbinom(100, 10, 0.3)==2)
r1000<-mean(rbinom(1000, 10, 0.3)==2)
r10000<-mean(rbinom(10000, 10, 0.3)==2)
r100000<-mean(rbinom(100000, 10, 0.3)==2)

r<-c(r10, r100, r1000, r10000, r100000)
n<-c(10, 100, 1000, 10000, 100000)

plot(n, r, xlab="Number of replicates", ylab="Probability of 70% flips coming up tails")
```

Produce a second plot but this time log-transform the number of trials.  Make sure to adjust your axis labels appropriately.  

```{r}
plot(log(n), r, xlab="Log of number of replicates", ylab="Probability of 70% flips coming up tails")
```

Q: Which simulation yields a result most similar to the exact probability?  What pattern do you see in your first graph?  What does yoru second graph reveal that your first graph did not?

A: The simulation with the largest sample size yields a result most similar to the exact probability.  The sample sizes differ by orders of magnitutde, which makes a log scale more appropriate for graphing.  Once you log-transform the number of replicates, you can see that sample sizes of 100 and larger start to converge on the same value (the exact probability, which you calculated in question 3 using dbinom(2, 10, 0.3)).



