---
title: "Probability exercises"
author: "SMWadgymar"
date: '2022-08-16'
output: html_document
---

####Exercises

1) Simulating coin flips:  Generate 10 separate random flips with probability 0.7 of producing heads.

```{r}

```

Q: What kind of values do you see and what do they represent?

A: 


2) Simulating draws from a binomial distribution:  Generate 100 occurrences of flipping 10 coins, each with 70% probability of producing heads. Produce a table of the probability distribution that you generate (hint: your table should report probabilities and not counts).

```{r}

```

Q: What kind of values do you produce and what do they represent?  Plot a histogram and describe the shape that your probability distribution takes.  What does the freq=F argument mean?

A: 
```{r}
hist(???, freq=F)
```



3) Calculate the exact probability that 2 heads will arise from 10 coin flips with a 70% probability of coming up tails.  

```{r}

```

Q: Compare your answer with a simulation of 10,000 trials.  Do the two approaches yield similar results?

A: 
```{r}

```




4) Calculate the cumulative probability that at least five coins out of 10 are heads with a 30% probability of coming up heads.  

```{r}

```

Q: Compare your answer with a simulation of 10,000 trials.  Do the two approaches yield similar results?

A: 
```{r}

```




5) Repeat the simulation you ran in exercise (3) with 10, 100, 1,000, 10,000, and 100,000 trials.  Produce a plot depicting the number of trails on the x-axis and the associated probabilities you calculated on the y-axis.  Produce a second plot but this time log-transform the number of trials.  Make sure to adjust your axis labels appropriately.  

```{r}

```

Q: Which simulation yields a result most similar to the exact probability?  What pattern do you see in your first graph?  What does yoru second graph reveal that your first graph did not?

A: 



####Background: joint events

6) Let's simulate two events, A and B.

```{r}
A<-rbinom(100000, 1, 0.5)
B<-rbinom(100000, 1, 0.5)
```

Q: What is the probability of event A occuring (that is, equaling 1)?  How would you calculate that in R?  What about the probability of event B occuring?

A:


Q: Using the General Addition Rule, estimate the probability of event A or event B occuring (show your notation).

A:


We can implement the General Addition Rule in R this way, where the | symbol means 'or':

```{r}
mean(A|B)
```


7) Q: Using the Multiplication Rule for Independent Events, estimate the probability of event A and event B occuring (show your notation).

A: 


We can implement the Multiplication Rule for Independent Events in R this way, where the & symbol means 'and':

```{r}
mean(A&B)
```


8) If events A and B are independent, and A has a 50% chance of happening and B has a 30% chance of happening, what is the probability that they will both happen?  Use simulations to answer this question.

```{r}

```


9) Expanding on exercise (19), event C has a 70% chance of happening.  What is the probability that events A, B, and C all happen?

```{r}

```


10) If events A and B are independent, and A has a 40% chance of coming up heads and B has a 75% chance of coming up heads, what is the probability that either A or B will come up heads?  Use simulations to answer this question.

```{r}

```


11) Suppose X is a random binomially distributed variable (10, 0.3) and Y is another random binomially distributed variable (10, 0.65), and that they are independent.  What is the probability that either of these variables is less than or equal to 5?  Estimate this probability both using simulations of 10,000 trials and by calculating exact cumulative densities.  

```{r}

```

Q: How do these two approaches compare?

A: 

