Random notes from our review session - I hope they make sense, but let me know if you have questions!

Rcode:

Finds p given a t-value
pt(t=#, df=, lower.tail=T/F)

gives you the cumulative probability/density for a given cut-off point using a normal distribution
pnorm(cutoff_point, mean, sd, lower.tail=T/F) #default to mean of 0 and sd of 1

gives you the cutoff point associated with a given probability
qnorm(probability, mean, sd, lower.tail=T/F)

example find the cutoff point associated with the value 0.2
qnorm(0.2, 0, 1) = -0.8416 (this is your cutoff point)

Confidence inetervals
point estimate +/- z* * SE
point estimate +/- t* * SE

Assumptions:
normality (shapiro test, qqplot, hist w/norm function)
independence (look for random)
equal variance across groups (var.test)


Single mean t-test
t.test(variable, mu = null#, alternative = "two.sided") 
two-sided is default, can also specify less or greater

Paired t-test
t.test(x1, x2, paired=TRUE, alternative="two.sided")

Two-mean t-test
t.test(x1, x2, paired=FALSE, alternative="two.sided")


Worked anova problem:

```{r}

caffeine <-data.frame(
ppm50=c(-0.4, 0.34, 0.19, 0.05, -0.14),
ppm100=c(0.01, -0.39, -0.08, -0.09, NA),
ppm150=c(0.65, 0.53, 0.39, -0.015, 0.46),
ppm200=c(0.24, 0.44, 0.13, 1.03, 0.05))
 
caffeine #view the datafile
 
caffeine_stack<-stack(caffeine)
names(caffeine_stack)<-c("values", "treatment")

#Y~X, response~explanatory
caffeine.aov<-aov(caffeine_stack$values~caffeine_stack$treatment)

summary(caffeine.aov)

#post-hoc, pair-wise testing
TukeyHSD(caffeine.aov)

```


Subsetting
```{r}
#Subsetting a dataset
VC0.5<-subset(ToothGrowth, supp=="VC" & dose==0.5)

VC<-subset(ToothGrowth, supp=="VC")
VC0.5<-subset(VC, dose==0.5)

#Subsetting a vector
ToothGrowth$len[ToothGrowth$dose==0.5]
```

